---
title: "Machine Learning Course Project"
author: "Alejandra Perez"
output: html_document
---

```{r setup, echo=TRUE}
# Set Working Directory
setwd("C:/Users/maria/OneDrive/Escritorio/R/Hopkins/Practical machine learning")
```

# Load necessary libraries
```{r load_libraries, echo=TRUE}
library(caret)
library(randomForest)
library(ggplot2)
```

# Load data
```{r load_data, echo=TRUE}
training <- read.csv("pml-training.csv", na.strings=c("NA","","#DIV/0!"))
testing <- read.csv("pml-testing.csv", na.strings=c("NA","","#DIV/0!"))
```

# Remove variables with too many missing values
```{r clean_data, echo=TRUE}
training <- training[, colSums(is.na(training)) == 0]
testing <- testing[, colSums(is.na(testing)) == 0]
```

# Remove first few identifier columns
```{r remove_identifiers, echo=TRUE}
training <- training[, -c(1:7)]
testing <- testing[, -c(1:7)]
```

# Convert target variable to factor
```{r convert_target, echo=TRUE}
training$classe <- as.factor(training$classe)
```

# Split data into training (75%) and validation (25%) sets
```{r split_data, echo=TRUE}
set.seed(123)
inTrain <- createDataPartition(training$classe, p = 0.75, list = FALSE)
trainData <- training[inTrain,]
valData <- training[-inTrain,]
```

# Train a Random Forest model
```{r train_model, echo=TRUE}
set.seed(123)
rf_model <- train(classe ~ ., data = trainData, method = "rf", 
                  trControl = trainControl(method="cv", number=5))
```

# Print model summary
```{r print_model, echo=TRUE}
print(rf_model)
```

# Predict on validation set
```{r predict_validation, echo=TRUE}
val_preds <- predict(rf_model, valData)
```

# Evaluate model performance
```{r evaluate_model, echo=TRUE}
conf_matrix <- confusionMatrix(val_preds, valData$classe)
print(conf_matrix)
```

# Predict on the test set
```{r predict_test, echo=TRUE}
final_preds <- predict(rf_model, testing)
```

# Save predictions for submission
```{r save_predictions, echo=TRUE}
write.csv(final_preds, file = "final_predictions.csv", row.names = FALSE)
```

# Print final message
```{r final_message, echo=TRUE}
cat("Machine Learning Project Completed!\n")
cat("Model: Random Forest\n")
cat("Final predictions saved as final_predictions.csv\n")
cat("Future improvements: Consider using Gradient Boosting or Neural Networks.\n")
